---
title: "Prediction Assignment Writeup"
author: "Aleksander Omelyanyuk"
date: "October 18, 2018"
output:
  word_document: default
  pdf_document: 
    fig_height: 7
  html_document: default
---

---

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, my goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

---

##What I should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

## Recieving data

All data in two tables pml-training.csv and pml-testing.csv. I will use pml-training data for model creating and testing. 
In code below I loading needable libraries and data.

```{r, reciewing data}
#### Libraries
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)

### Data loading
urlTrain<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_final_test<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
my_data<-read.csv(urlTrain,na.strings = c("NA","#DIV/0!",""))
final_test<-read.csv(url_final_test,na.strings = c("NA","#DIV/0!",""))

###loaded data
str(my_data)

```

---

##Preprocessing the Data

As wee can see, in tables with data many fields with "NA" variables. I must clean data for analysis. In code below I well do it.
Some variables have no sense, like "X", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", I remove this variables too.
For training and testing I will use 75 and 25 percents of data "pml-training.csv".

```{r, preprocessing}
cut_off <- apply(my_data, 2, function(x) sum(is.na(x)))/nrow(my_data)

my_data <- my_data[!(cut_off > .95)]
final_test<-final_test[!(cut_off > .95)]

cut_off<-nearZeroVar(my_data)

my_data<-my_data[,-cut_off]
final_test<-final_test[,-cut_off]

my_data<-my_data[,-c(1,3,4,5)]
final_test<-final_test[,-c(1,3,4,5)]

str(my_data)

my_train<-createDataPartition(my_data$classe,p=.75,list = FALSE)

train_data<-my_data[my_train,]
test_data<-my_data[-my_train,]
```

---

##Processing the Data

### Rpart model

```{r, processing rpart model}
set.seed(1)
fit_rpart<-rpart(data = train_data,classe ~ ., method = "class")
pred_rpart<-predict(fit_rpart,test_data,type = "class")

rpart_prediction<-confusionMatrix(pred_rpart,test_data$classe)
print(rpart_prediction)

varImp(fit_rpart)


```

We can see not all variables take part in model.

---

###Random Forest model

```{r, processing random forest model}
set.seed(1)
fit_rf<-randomForest(data = train_data, classe ~ ., method = "class")
pred_rf<-predict(fit_rf,test_data, type = "class")
rf_prediction<-confusionMatrix(pred_rf,test_data$classe)
print(rf_prediction)

varImpPlot(fit_rf)
```

This model more slower, but more accuracy, in my mind all variables take part in model creation.

---

###Final result

For final result I use random forest model, because it more accuracy.

```{r, final processing}
final_predict<-predict(fit_rf,final_test, method = "class")
print(final_predict)
```

## RESULT
In this work I use 2 models random forest and rpart.
Rpart more faster but random forest have more accuracy.
